---
title: "STAT4100 HW2"
author: "Cody Frisby"
date: "January 14, 2016"
output: pdf_document
---

### #3

* First I will bring the data into R, plot it, and run a quick t test.

```{r}
dat <- data.frame(treatment = c("A", "B", "A", "A", "B"), 
                  response = c(14, 16, 19, 17, 15))
boxplot(dat$response ~ dat$treatment, col = "lightblue")
t <- t.test(response ~ treatment, data = dat, var.equal = TRUE, alternative = "less")
```

Do not reject null hypothesis.  P value = `r t$p.value`.

### #4 Two procedures for predicting shear strength will be compared.

#### a) Is there evidence to support the claim that there is a differnce between the two methods?

```{r}
Karl <- c(1.186, 1.151, 1.322, 1.339, 1.2, 1.402, 1.365, 1.537, 1.559)
Lehigh <- c(1.061, 0.992, 1.063, 1.062, 1.065, 1.178, 1.037, 1.086, 1.052)
s <- t.test(Karl, Lehigh, paired = TRUE)
boxplot(Karl, Lehigh, col = "lightblue")
s
```


Yes, there is evidence to suggest a difference in mean performance     
between the two methods at $\alpha$ = 0.05.

#### b) The p value for this test is
$\ `r s$p.value`$.  This is much smaller than 0.05.

#### c) Confidence Interval:
[`r s$conf.int`].  Zero is not contained in this interval.

#### d) Normality Assumptions:

* QQ Plot for the Karl and Lehigh data.  

```{r}
q1 <- qqnorm(Karl, plot.it = FALSE)
q2 <- qqnorm(Lehigh, plot.it = FALSE)
plot(range(q1$x, q2$x), range(q1$y, q2$y), type = "n", xlab = "Theoretical Quantiles", ylab = "Sample Quantiles", main = "Normal Q-Q Plot")
points(q1, pch = 16)
points(q2, col = "red", pch = 3)
qqline(Karl); qqline(Lehigh)
```

The variances of the two samples is not equal.  The var.test function in R can show this.  The difference in the two lines also demonstrates this. 

```{r}
var.test(Karl, Lehigh)
```


#### e) Normality for difference in ratios:

* Here we plot the difference between the two variables and run a shapiro.wilk test.

```{r}
qqnorm(Karl - Lehigh)
qqline(Karl - Lehigh)
shapiro.test(Karl - Lehigh)
```

Looks like we can reasonably assume normality for the difference between the two methods.  

#### f) Discuss the role of the normality assumption in the paired t test.
* The normality assumption is only moderately important with a paired t test just like other t tests.  With a paired t test the assumption of normality applies to the differences.  The individual sample measurements do not have to be normally distributed.

### #5

```{r}
y1 <- 12.5
y2 <- 10.2
s1 <- 101.17
s2 <- 94.73
n1 <- 8
n2 <- 9
# compute the F statistic for 7,8 degrees of freedom at alpha=0.05
qf(0.975, 7, 8)
s1/s2
# is the ratio of the two variances greater?
s1/s2 > qf(0.975, 7, 8)
```

#### a)
No. `r s1/s2` is not greater than `r qf(0.975, 7, 8)`.  We cannot conclude that the variances are different.  

```{r}
sp <- sqrt(((n1-1)*s1 + (n2-1)*s2) / (n1 + n2 - 2))
t0 <- (y1 - y2) / (sp * sqrt((1/n1) + (1/n2)))
qt(0.95, 15)
t0
t0 > qt(0.95, 15)
```

#### b)
Do not reject.  There is not evidence to indicate the new filtering device has affected the mean.